---
title: "Using xtracks"
author: "Brian Wood"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using xtracks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Installing xtracks

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("brianwood1/xtracks")
```

### Load xtracks

```{r}
library(xtracks)
```

### load some trackpoint / spatiotemporal data

First you need to load some trackpoint data into a dataframe. For demonstration purposes, the xtracks package comes with two dataframes named <code>d1</code> and <code>d2</code> that are appropriately formatted. The structure of the data can be seen below. 

```{r, load_data, echo=TRUE}
head(d1)
```

As you can see, the data xtracks objects are based on includes the sort that come from GPS devices (lat, lon, elevation, time stamp, elevation) but also include a 'distance from camp' measure for all trackpoints and an 'in camp' flag. These later values are critical for analyses performed by xtracks and should be computed as an initial data processing step. 

### construct an xtracks object

To construct an xtrack object, one must specify:

- lat
- lon
- elevation
- in-camp status
- time
- distance from camp centroid
- whether each trackpoint is "in camp" or not
- the utm_epsg code. 

lat and lon are expected to be in decimal-degree, WGS 84 format, which is the default in most GPS devices mobile devices. elevation is expected to be in meters above sea level. The "in_camp" parameter refers to whether each trackpoint is within or outside the boundaries of a residential area, which in Wood et al. 2021 refers to the spatial boundaries of a Hadza camp; but could more generally be considered the boundaries of a residential or habitation area, something like a village or a camp, as appropriate in a given field setting. This is useful for indicating travel for the purpose of aquiring resources -- AKA foraging travel, and needed for sinuosity measures. Distance from camp centroid is expected to be the as-the-crow-flies distance from the center of a residential area / 'camp' in meters (also needed for sinuosity measures). The epsg code identifies the UTM zone of your study location. This is needed for projecting lat / lon coordinates into UTM space. To find the epsg code for your study location region of your track, check out https://spatialreference.org/ref/epsg/

Assuming these initial data are in place, the code to construct an xtrack object is as follows:

```{r, construct_xtracks}
xt_1 <- xtrack(lat=d1$lat, lon=d1$lon, elevation_m=d1$elevation_m, in_camp=d1$in_camp, unix_time=d1$unix_time, distance_from_camp_m=d1$distance_from_camp_m, utm_epsg=32736)
xt_2 <- xtrack(lat=d2$lat, lon=d2$lon, elevation_m=d2$elevation_m, in_camp=d2$in_camp, unix_time=d2$unix_time, distance_from_camp_m=d2$distance_from_camp_m, utm_epsg=32736)
```

